Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-11-18 appendix.tex
sup

In this supplementary material, we provide further technical details and the
less important discussions that are not included in the manuscript due to space
limit.
%
Discussions required by the author guidelines are also included in this
supplementary material.

@@

We adopt ResNet-18 following the state-of-the-art defense~\cite{robrank} for fair comparison.
%
Different backbone models are expected to achieve significantly different
robustness and benign example performance.

@@

\subsection{State-of-The-Art Defenses}
\label{sec:b1}

@@

The direct formulation of min-max adversarial training~\cite{madry} for
triplet loss-based DML is:
%
\begin{equation}
%
	\theta^* = \argmin_\theta [ \argmax_{\vr_a, \vr_p, \vr_n} L_\text{T}(
	\mA + \vr_a, \mP + \vr_p, \mN + \vr_n)]
%
\end{equation}
%
Previous works~\cite{advrank,robrank} point out this method will easily lead
to model collapse.
%
Our observation is the same.

@@

~\newline
\noindent\textbf{II. Range of Hardness.}

@@

~\newline
\noindent\textbf{III. Mixing HM[$\mathcal{R},\mathcal{M}$] and
HM[$\mathcal{S},\mathcal{M}$]}

According to the experimental results, HM[$\mathcal{R},\mathcal{M}$] achieves 
a high ERS and adversarial training efficiency, but suffers from a low R@1.
%
HM[$\mathcal{S},\mathcal{M}$] achieves a high R@1, but its ERS and efficiency
could be further improved.
%
They can be mixed for adversarial training with a probability $p$ to combine
their advantages, but the eventual overall performance is still lower than
that of HM[$\mathcal{S},g_\mathsf{LGA}$] according to our observation.

@@

\oo{TODO}

@@

\noindent\textbf{I. Parameters or Adjustable Items}

The ICS loss term can be appended to the loss for adversarial training.
%
The only parameter for ICS loss term is the weight constant $\lambda$.
%
The selection of this parameter is discussed in the previous section of this
supplementary material (see ``Performance-Sensitive Factors'').

@@

According to \cite{robrank}, when the embedding vectors are normalized onto the
real unit hypersphere and Euclidean distance is used, the gradients of the
triplet loss with respect to the anchor, positive, and negative embedding
vectors are respectively:
%
\begin{align}
%
	\frac{\partial L_\text{T}}{\partial \va} &= \frac{\va - \vp}{\|\va - \vp\|}
	- \frac{\va - \vn}{\| \va - \vn \|} \\
	\frac{\partial L_\text{T}}{\partial \vp} &= \frac{\vp - \va}{\|\va - \vp\|} \\
	\frac{\partial L_\text{T}}{\partial \vn} &= \frac{\va - \vn}{\|\va - \vn\|},
%
\end{align}
%
when $L_\text{T}>0$.
And the above equations have been reflected in Fig.~5 (a) with correct directions.

@@

Let $\tilde{\va}$, $\tilde{\vp}$, and $\tilde{\vn}$ be $\phi(\mA+\hat{\vr}_a)$,
$\phi(\mP+\hat{\vr}_p)$ and $\phi(\mN+\hat{\vr}_n)$ respectively.
%
In our proposed ICS loss term, only $(\va, \tilde{\va}, \vp)$ are involved.
%
Other alternative sections of triplets from the sextuplet are also
possible.

\ul{(1) Why not use $L_\text{T}(\va, \tilde{\va}, \tilde{\vp})$? }

As shown in Fig.~5 (c) in the manuscript, the position of $\tilde{\vp}$ is
always further away from both $\va$ and $\tilde{\va}$ than $\vp$ due to
the gradient direction.
%
Thus, the loss term in the question is not as effective as $L_\text{ICS}$.

\ul{(2) What if $L_\text{T}(\va, \vp, \vn)$ is used? Namely, the regular training
and adversarial training are mixed together.}

According to our observation, mixing regular training and adversarial training
leads to better R@1 but very drastic robustness degradation for both ACT and
our proposed defense.

\ul{(3) Why not use the symmetric counterpart of $L_\text{ICS}$, \ie,
		$L_\text{T}(\vp, \tilde{\vp}, \va)$?}

Every sample in the training dataset will be used as anchor for once per
epoch.
%
Such symmetric loss term is duplication to $L_\text{ICS}$ and is
not necessary.
%
Experimental results suggest negligible difference.

\ul{(4) $\tilde{\vn}$ is very close to $\va$ in Fig.~5 in manuscript.
Why not use $L_\text{T}(\va, \tilde{\va}, \tilde{\vn})$?}

The loss term in question is enforcing inter-class structure instead of
intra-class structure.
%
Besides, experimental results suggest negligible difference.
%
We speculate this loss term is duplicated to the adversarial training term,
\ie, $L_\text{T}(\tilde{\va}, \tilde{\vp}, \tilde{\vn})$, which enforces
inter-class structure in a stronger manner.

\ul{(5) Similar to question (4), why not use $L_\text{T}(\va, \vp, \tilde{\vn})$ ?}

@@

\subsection{Empirical Robustness Score (ERS)}

\subsection{Hardware and Software Platform}

We conduct the experiments with two Nvidia Titan Xp GPUs (12GB of memory each)
under the Ubuntu 16.04 operating system with PyTorch 1.8.2 in distributed data
parallel mode.
%
Note, the training result of all methods in distributed data parallel mode
slightly differs from the result obtained in single-GPU mode, due to the
asynchronized gradient updates in the distributed mode (this is a common
phenomenon in distributed deep learning).
%
In fact, our method can be even more efficient in single-GPU mode thanks to the
synchornized gradients, but single-GPU mode will be extremely time consuming
(more than a week for a single run of the most time-consuming experiment).
%
All of our experiments are conducted in distributed data parallel mode for ease
of comparison, as well as controlling the time cost.

@@

>>>>

In this supplementary material, we provide more technical details and
additional discussions that are excluded from the manuscript due to space
limit.

@@

We adopt ResNet-18 following the state-of-the-art defense~\cite{robrank} for
fair comparison.
%
Since our proposed method is independent to the backbone choice, and hence
is expected to be effective with different backbone models.

\oo{TODO: mnasnet ghmetsm ghmetsmi}

@@

\subsection{Difference between Existing Defenses \& HM}
\label{sec:b1}

@@

The direct formulation of min-max adversarial training~\cite{madry} for
triplet loss-based DML is:
%
\begin{equation}
%
	\theta^* = \argmin_\theta [ \argmax_{\vr_a, \vr_p, \vr_n} L_\text{T}(
	\mA + \vr_a, \mP + \vr_p, \mN + \vr_n)]
%
\end{equation}
%
Previous works~\cite{advrank,robrank} point out this method will easily lead
to model collapse.
%
Our observation suggests the same.

@@

~\newline
\noindent\textbf{II. Extreme Values of Hardness.}

@@

@@

@@

The ICS loss term can be appended to the loss for adversarial training.
%
The only parameter for ICS loss term is the weight constant $\lambda$,
and has been discussed in the ``Performance-Sensitive Factors'' part
of the previous section.

@@

According to \cite{robrank}, when the embedding vectors are normalized onto the
real unit hypersphere and Euclidean distance is used, the gradients of the
triplet loss with respect to the anchor, positive, and negative embedding
vectors are respectively:
%
\begin{align}
%
	\frac{\partial L_\text{T}}{\partial \va} &= \frac{\va - \vp}{\|\va - \vp\|}
	- \frac{\va - \vn}{\| \va - \vn \|} \\
	\frac{\partial L_\text{T}}{\partial \vp} &= \frac{\vp - \va}{\|\va - \vp\|} \\
	\frac{\partial L_\text{T}}{\partial \vn} &= \frac{\va - \vn}{\|\va - \vn\|},
%
\end{align}
%
when $L_\text{T}>0$.
And the above equations have been reflected in Fig.~5 (a) in terms of vector direction.

@@

Let $\tilde{\va}$, $\tilde{\vp}$, and $\tilde{\vn}$ be $\phi(\mA+\hat{\vr}_a)$,
$\phi(\mP+\hat{\vr}_p)$ and $\phi(\mN+\hat{\vr}_n)$ respectively.
%
In our proposed ICS loss term, only $(\va, \tilde{\va}, \vp)$ are involved.
%
Other alternative selections of triplets from the sextuplet are possible,
but are not as effective as $L_\text{ICS}$ for improving adversarial robustness.

\ul{(1) $L_\text{T}(\va, \tilde{\va}, \tilde{\vp})$}

As shown in Fig.~5 (c) in the manuscript, the position of $\tilde{\vp}$ is
always further away from both $\va$ and $\tilde{\va}$ than $\vp$ due to
the gradient direction.
%
Thus, the loss value of $L_\text{T}(\va, \tilde{\va}, \tilde{\vp})$ will always
be smaller than $L_\text{ICS}$, and hence is less effective than $L_\text{ICS}$.

\ul{(2) $L_\text{T}(\va, \vp, \vn)$: Mixing regular training
and adversarial training.}

According to our observation, mixing regular training and adversarial training
leads to better R@1 with drastic robustness degradation for both ACT and
our defense.

\ul{(3) $L_\text{T}(\vp, \tilde{\vp}, \va)$: Symmetric counterpart of $L_\text{ICS}$.}

Every sample in the training dataset will be used as anchor for once per
epoch.
%
Such symmetric loss term is duplicated to $L_\text{ICS}$ and is
not necessary.
%
Experimental results suggest negligible difference compared to $L_\text{ICS}$.

\ul{(4) $L_\text{T}(\va, \tilde{\va}, \tilde{\vn})$: $\tilde{\vn}$ is very close to $\va$ in Fig.~5.}

It enforces \emph{inter}-class structure instead of intra-class structure.
%
Besides, experimental results suggest negligible difference.
%
We speculate this loss term is duplicated to the adversarial training loss term,
\ie, $L_\text{T}(\tilde{\va}, \tilde{\vp}, \tilde{\vn})$, which enforces
inter-class structure as well in a stronger manner.

\ul{(5) $L_\text{T}(\va, \vp, \tilde{\vn})$}

@@

@@

\noindent\textbf{I Hardware and Software Configuration}

We conduct the experiments with two Nvidia Titan Xp GPUs (12GB of memory each)
under the Ubuntu 16.04 operating system with PyTorch 1.8.2 in distributed data
parallel.
